Fitting metabolic models to dissolved oxygen data: The Estuarine BAyesian Single-station Estimation method

### 2023-07-28 manu-draft.qmd
text edits for default prior equations

The remaining three parameters $a$ (light efficiency, $\left(\text{mmol}/\text{m}^2/\text{d}\right)/\left(\text{W}/\text{m}^2\right)$), $R$ (respiration, $\text{mmol}/\text{m}^3/\text{d}$), and $b$ ($\left(\text{cm}/\text{hr}\right)/\left(\text{m}^2/\text{s}^2\right)$) are estimated by likelihood given the observed data and prior distributions for the parameters. The JAGS software [as for BASE, @Plummer03] is used with the EBASE R package [described below, @Beck23; @RCT23] to estimate the unknown parameters using Markov Chain Monte Carlo (MCMC) simulations in a Bayesian framework. At each time step, the change in oxygen concentration between time steps is calculated from the equation using model inputs and parameter guesses based on the prior distributions, and then a finite difference approximation is used to estimate modeled oxygen concentration. Estimations proceed using a forward in time integration where the dissolved oxygen at each time step is informed by parameter estimates from the preceding time steps. The estimated dissolved oxygen concentration is also returned, which can be compared to the observed values as one measure of model performance. All parameter inputs and outputs for EBASE are shown in `r run_reference(id = "tbl-ebaseiotab")`. Areal rates for $P$, $R$, and $D$ ($\text{mmol}/\text{m}^2/\text{d}$) are returned for comparability of estimates across locations where water column depth may vary.

@@

$$
a \sim N(0.2, 1)T(0,\infty) \;\; \left(\text{mmol}/\text{m}^2/\text{d}\right)/\left(\text{W}/\text{m}^2\right)
$$ {#eq-aprior}

$$
R \sim N(20, 50)T(0,\infty) \;\; \text{mmol}/\text{m}^3/\text{d}
$$ {#eq-rprior}

>>>>

The remaining three parameters $a$ (light efficiency, $\left(\text{mmol}/\text{m}^2/\text{d}\right)/\left(\text{W}/\text{m}^2\right)$), $R$ (respiration, $\text{mmol}/\text{m}^3/\text{d}$), and $b$ ($\left(\text{cm}/\text{hr}\right)/\left(\text{m}^2/\text{s}^2\right)$) are estimated by likelihood given the observed data and prior distributions for the parameters. The JAGS software [as for BASE, @Plummer03] is used with the EBASE R package [described below, @Beck23; @RCT23] to estimate the unknown parameters using Markov Chain Monte Carlo (MCMC) simulations in a Bayesian framework. At each time step, the change in oxygen concentration between time steps is calculated from the equation using model inputs and parameter guesses based on the prior distributions, and then a finite difference approximation is used to estimate modeled oxygen concentration. Estimations proceed using a forward-in-time integration where the dissolved oxygen at each time step is informed by parameter estimates from the preceding time steps. The estimated dissolved oxygen concentration is also returned, which can be compared to the observed values as one measure of model performance. All parameter inputs and outputs for EBASE are shown in `r run_reference(id = "tbl-ebaseiotab")`. Areal rates for $P$, $R$, and $D$ ($\text{mmol}/\text{m}^2/\text{d}$) are returned for comparability of estimates across locations where water column depth may vary.

@@

$$
a \sim N(4, 1)T(0,\infty) \;\; \left(\text{mmol}/\text{m}^2/\text{d}\right)/\left(\text{W}/\text{m}^2\right)
$$ {#eq-aprior}

$$
R \sim N(300, 50)T(0,\infty) \;\; \text{mmol}/\text{m}^2/\text{d}
$$ {#eq-rprior}
